---
title: "Factors Affecting Miles Per Gallon"
author: "Yilmaz Rona"
date: "December 12, 2015"
output: html_document
---

# Introduction

We analyze [data extracted from the 1974 Motor Trend magazine describing fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models)](http://www.mortality.org/INdb/2008/02/12/8/document.pdf).

We will construct a model that predicts the fuel efficiency in miles per gallon (mpg) for these vehicles.

# Single variable correlations

```{r  cache=FALSE,echo=FALSE}
knitr::read_chunk('singlecorrelations.r')
```
```{r singlecorrelations, echo=FALSE}
```
```{r calcSingleCorr}
  single.correlations <- singlecorrelations()
  single.correlations
```

# Multivariate model

## Iteration 1

## Iteration 2

## Iteration 3


# Appendix 1: Data and Methods

We used R to perform the analysis.  The dataset is provided by default with the R installation.

## Data

```{r dataset, echo=FALSE}
mtcars

```

## Figures

```{r  cache=FALSE,echo=FALSE}
knitr::read_chunk('plot01.r')
```
```{r plot01, echo=FALSE}
```
```{r displayplot01,fig.height=8}
  plot01.predictors <- as.character( single.correlations$Pred)
  plot01(plot01.predictors)
```

## Code

```{r singlecorrelations, eval=FALSE}
```