---
title: "Fuel Efficiency and Transmission Types"
author: "Yilmaz Rona"
date: "December 12, 2015"
output: html_document
---

# Executive Summary

# Introduction

We analyze [data extracted from the 1974 Motor Trend magazine describing fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models)](http://www.mortality.org/INdb/2008/02/12/8/document.pdf) to explore the following:

* Question: Is an automatic or manual transmission better for MPG?
* Quantify: The MPG difference between automatic and manual transmissions.
* Explore: The relationship between miles per gallon and the other parameters in the data set.

# The Relationship Between Fuel Efficiency and Other Parameters

We begin by exploring the general relationship between MPG and the other parameters in the data.  
```{r  cache=FALSE,echo=FALSE}
knitr::read_chunk('singlecorrelations.r')
```
```{r singlecorrelations, echo=FALSE}
```
```{r calcSingleCorr}
  single.correlations <- singlecorrelations()
  single.correlations
```

# Transmission Type and Fuel Efficiency

## Iteration 1

## Iteration 2

## Iteration 3


# Appendix 1: Figures

## Figure 1: Fuel Efficiency (mpg) vs other parameters

```{r  cache=FALSE,echo=FALSE}
knitr::read_chunk('plot01.r')
```
```{r plot01, echo=FALSE}
```
```{r displayplot01,fig.height=8}
  plot01.predictors <- as.character( single.correlations$Pred)
  plot01(plot01.predictors)
```